---
title: "Non-Imaging (Overview)"
---

```{r}
#| label: setup
#| include: false 

for (file in list.files("../../../R", full.names = TRUE)) source(file)

```

<!-- Invisible reference to the R logo so the resource can be captured by quarto -->
<div style="display: none;">
[](/assets/img/documentation/rlogo.svg)
</div>

Following is a list of the ABCD non-imaging tables, across the domains. Users can also find the tables in a particular year via sorting by clicking on the specified Events column heading. For more detailed information, please refer to the respective domain sections.    

:::{.callout-note title="Imaging data"}
Corresponding imaging data are collected on even years. For more information, please refer to the [Imaging section](../imaging/index.qmd). 
:::

# Sort by Domain

::: {.panel-tabset}

### All
```{r}
create_table_info("domain_abbr == 'ab' | domain_abbr == 'fc' | domain_abbr == 'gn' | domain_abbr == 'le' | domain_abbr == 'mh' | domain_abbr == 'nc' | domain_abbr == 'nt' | domain_abbr == 'ph' | domain_abbr == 'su'")
```

### General
```{r}
create_table_info("domain_abbr == 'ab'")
```

### Friends, Family, & Community
```{r}
create_table_info("domain_abbr == 'fc'")
```

### Genetics
```{r}
create_table_info("domain_abbr == 'gn'")
```

### Linked External Data
```{r}
create_table_info("domain_abbr == 'le'")
```

### Mental Health
```{r}
create_table_info("domain_abbr == 'mh'")
```

### Neurocognition
```{r}
create_table_info("domain_abbr == 'nc'")
```

### Novel Technologies
```{r}
create_table_info("domain_abbr == 'nt'")
```

### Physical Health
```{r}
create_table_info("domain_abbr == 'ph'")
```

### Substance Use
```{r}
create_table_info("domain_abbr == 'su'")
```

:::


